---
title: "Occurrence of quinolone resistant *E. coli* originating from different animal species in Norway R Script"
output:
  pdf_document: default
  html_notebook: default
  html_document: default
---

```{r setup, include=TRUE, message=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, include = TRUE)
```

# Information
This script prepares all tables and figures and calculates necessary statistics from the relevant datasets.

The authors would like to thank the Stack Overflow community at www.stackoverflow.com for their help
with this script.

# Packages
```{r}
library(rlang)
library(tidyverse)
library(broom)
library(pander)
library(ggsci)
```

#Imports datasets and sets fonts used for figures

```{r}
# Main dataset
Dataset1 <-read.table(paste0("Dataset1.txt"),header = T, sep="\t",
                             stringsAsFactors = FALSE)

# Dataset for isolates with municipality information
Dataset2 <-read.table(paste0("Dataset2.txt"),header = T, sep="\t",
                             stringsAsFactors = FALSE)

# Sets fonts used in figures
windowsFonts(Times=windowsFont("TT Times New Roman"))
```

\newpage
# Functions
## Functions used in this analysis

```{r}
# Calculates 95 % confidence intervals
get_binCI <- function(x, n) as.numeric(setNames(binom.test(x,n)$conf.int*100,
                                                c("lwr", "upr")))

# Renames column values to full species names
fullnames<-function(dataset, ..., originals=c("can","chi","chila","bov",
                                   "fox","avi","sau","equ",
                                   "ren","pig","tur"),
                     newnames=c("Dogs","Broilers","Layers",
                                "Cattle","Red Foxes","Wild Birds",
                                "Sheep","Horses","Reindeer",
                                "Pigs","Turkeys")){
  names(newnames) = originals
  dots = quos(...)
  dataset %>% 
    mutate_at(vars(!!! dots), as.character) %>%
    mutate_at(vars(!!! dots), funs(newnames[.])) %>%
    data.frame(row.names = NULL)
}


# MIC distribution tables, per MIC per year
MIC_dist <- function(dataset, ...){
  dataset %>%
  count(..., year) %>%
  mutate(year = as.numeric(year)) %>%
  group_by(year) %>%
  mutate(perc_MIC = n/sum(n)*100) %>%
  spread(..., perc_MIC) %>%
  aggregate(., by = list(.$year), na.rm = TRUE, FUN = sum) %>%
  select(-year) %>%
  rename(year = Group.1)
}

```
\newpage
#MIC distribution tables

```{r kable}
# Table for CIP MIC distribution per year
MIC_CIP_year <- MIC_dist(Dataset1, CIP)
MIC_CIP_year <- MIC_CIP_year %>%
  mutate(sum_res = rowSums(.[6:11]))
MIC_CIP_year[,2:14] <- round(MIC_CIP_year[,2:14],1)
pander(MIC_CIP_year)

# Table for NAL MIC distribution per year
MIC_NAL_year <- MIC_dist(Dataset1, NAL)
MIC_NAL_year <- MIC_NAL_year %>%
  mutate(sum_res = rowSums(.[7:10]))
MIC_NAL_year[,2:12] <- round(MIC_NAL_year[,2:12],1)
pander(MIC_NAL_year)

# Table for percent quinolone resistance per year, CI and 
# number of isolates for CIP MICs
perc_year_CIP <- Dataset1 %>%
  mutate(CIP_res = ifelse(CIP > 0.06,
                          "Resistant",
                          "NonResistant")) %>%
  count(CIP_res, year) %>%
  group_by(year) %>%
  spread(CIP_res, n , fill = 0) %>%
  mutate(n = Resistant + NonResistant,
         perc_res = Resistant/n*100,
         lwr = get_binCI(Resistant, n)[1],
         upr = get_binCI(Resistant, n)[2]) %>%
  select(year,Resistant,n,perc_res,lwr,upr) %>%
  rename(Year=year,
         "Resistance(%)"=perc_res,
         "95 % CI Low"=lwr,
         "95 % CI High"=upr)

perc_year_CIP[,4:6] <- round(perc_year_CIP[,4:6],1)

pander(perc_year_CIP)

# Table for percent quinolone resistance per year, CI and 
# number of isolates for NAL MICs
perc_year_NAL <- Dataset1 %>%
  mutate(NAL_res = ifelse(NAL > 16,
                          "Resistant",
                          "NonResistant")) %>%
  count(NAL_res, year) %>%
  group_by(year) %>%
  spread(NAL_res, n , fill = 0) %>%
  mutate(n = Resistant + NonResistant,
         perc_res = Resistant/n*100,
         lwr = get_binCI(Resistant, n)[1],
         upr = get_binCI(Resistant, n)[2]) %>%
  select(year,Resistant,n,perc_res,lwr,upr) %>%
  rename(Year=year,
         "Resistance(%)"=perc_res,
         "95 % CI Low"=lwr,
         "95 % CI High"=upr)

perc_year_NAL[,4:6] <- round(perc_year_NAL[,4:6],1)

pander(perc_year_NAL)

# MIC distribution for various antibiotics
# (only for the resistant isolates)
order_vector <- c("TET","CHL","AMP","CTX",
                  "SMX","TMP","GEN","CIP","NAL")

MIC_dist_AB <- Dataset1 %>%
  filter(quin_res==1) %>%
  select(.,AMP,TET,CHL,SMX,TMP,GEN,CTX,CIP,NAL) %>%
  gather() %>%
  group_by(key) %>%
  mutate(n1 = n()) %>%
  count(value,n1) %>%
  mutate(perc = n/sum(n)*100) %>%
  select(-n,n=n1) %>%
  spread(value,perc,fill=0) %>%
  ungroup() %>%
  slice(match(order_vector, key)) %>%
  rename(Antimicrobial = key,
         ">256" = `2048`)

MIC_dist_AB[,3:17] <- round(MIC_dist_AB[,3:17],1)

pander(MIC_dist_AB)

# Table for percent resistance per antibiotic, CI, 
# and number of isolates for the QREC isolates
perc_AB <- Dataset1 %>%
  filter(quin_res == 1) %>%
  select(amp_res, tet_res, chl_res, smx_res, 
         tmp_res, gen_res, ctx_res, cip_res, nal_res) %>%
  gather() %>%
  group_by(key) %>%
  mutate(n1 = n()) %>%
  count(value, n1) %>%
  mutate(value = ifelse(value == 1,
                        "Resistant",
                        "NonResistant"),
         Antibiotic = key,
         Antibiotic = toupper(gsub("_res", "", key))) %>%
  spread(value, n , fill = 0) %>%
  rename(n = n1) %>%
  mutate(perc_res = Resistant/n*100,
         lwr = get_binCI(Resistant, n)[1],
         upr = get_binCI(Resistant, n)[2]) %>%
  ungroup() %>%
  select(Antibiotic,Resistant,n,perc_res,lwr,upr) %>%
  rename("Resistance(%)"=perc_res,
         "95 % CI Low"=lwr,
         "95 % CI High"=upr) %>%
  slice(match(order_vector, Antibiotic))

perc_AB[,4:6] <- round(perc_AB[,4:6],1)

pander(perc_AB)

# MIC values for the two QREC isolates with sub ECOFF
# NAL MIC
high_CIP_low_NAL <- Dataset1 %>%
  filter(CIP > 0.12 & NAL < 16) %>%
  select(CIP, NAL)

pander(high_CIP_low_NAL)

# MIC values for the two QREC isolates with sub ECOFF
# CIP MIC
low_CIP_high_NAL <- Dataset1 %>%
  filter(CIP < 0.12 & NAL > 16) %>%
  select(CIP, NAL)

pander(low_CIP_high_NAL)
```
\newpage
#Detection of resistance to multiple antimicrobials

```{r}
# Creates a table with number of co-resistant 
# isolates and percentage of the resistant isolates
multiResTable <- Dataset1 %>%
  count(quin_res,amp_res,tet_res,chl_res,
        smx_res,tmp_res,gen_res,ctx_res) %>%
  filter(quin_res == 1) %>%
  mutate(multi_res = rowSums(.[1:8])) %>%
  select(n,multi_res) %>%
  aggregate(n~multi_res,.,sum) %>%
  mutate(pIso = n/sum(n)*100) %>%
  rename("No. of Antimicrobials"=multi_res,
         "No. of Isolates"=n, "Percent of Isolates"=pIso)

pander(multiResTable)
```
\newpage
#CIP/NAL MIC comparisons

```{r}
# Prepares the table with counts of all combinations
# of CIP/NAL MICs for the QREC isolates
MIC_CIP_NAL_comp <- Dataset1 %>%
  filter(quin_res == 1) %>%
  count(CIP,NAL) %>%
  mutate(CIP = ifelse(CIP == 16, ">8",
               ifelse(CIP == 0.016, 0.015, CIP)),
         NAL = ifelse(NAL == 256, ">128",NAL))

# Creates a plot of the table above
ggplot(MIC_CIP_NAL_comp,aes(factor(CIP,levels=unique(CIP)),
                                factor(NAL,levels=c("1","2","4",
                                                    "8","16","32",
                                                    "64","128",
                                                    ">128")),
                                    size=factor(n)))+
  geom_point()+
  geom_vline(xintercept = 2.5, color = "#999999", 
             size = 0.3, alpha = 0.8)+
  geom_hline(yintercept = 2.5, color = "#999999", 
             size = 0.3, alpha = 0.8)+
  geom_rect(data = data.frame(x = c(1:8)), 
            aes(xmin = x - 0.5, xmax = x + 0.5,
                ymin = 0.42, ymax = 0.5),
            col = "#666666",
            fill = NA,
            inherit.aes = F)+
  geom_rect(data = data.frame(y = c(1:6)), 
            aes(xmin = 0.42, xmax = 0.5,
                ymin = y - 0.5, ymax = y + 0.5),
            col = "#666666",
            fill = NA,
            inherit.aes = F)+
  theme(panel.grid = element_blank(),
        panel.background = element_blank(),
        axis.ticks = element_blank(),
        legend.key = element_blank(),
        text = element_text(family="Times"))+
  labs(x="CIP MIC Levels",
       y="NAL MIC Levels",
       size="Number of\nisolates")
```
\newpage

#Species data

##Creates a table with percent resistance per species, with confidence intervals

```{r}
# Creates a table with information on the species
species_information <- Dataset1 %>%
  count(NORMart,quin_res) %>%
  mutate(quin_res = ifelse(quin_res == 0,
                           "NonResistant",
                           "Resistant")) %>%
  spread(quin_res,n) %>%
  mutate(Resistant = ifelse(is.na(Resistant),0, Resistant),
         n = NonResistant + Resistant,
         pRes = Resistant/n*100) %>%
  group_by(Resistant, n) %>%
  mutate(lwr = get_binCI(Resistant, n)[1],
         upr = get_binCI(Resistant, n)[2]) %>%
  fullnames(., NORMart) %>%
  select(NORMart, n,Resistant, pRes,lwr,upr) %>%
  rename(Species = NORMart, "Resistance (%)" = pRes,
         "95 % CI Low" = lwr, "95 % CI High" = upr)
species_information[,4:6] <- round(species_information[,4:6],1)
pander(species_information)

```
\newpage

#Broiler data

##Creates a table with information about resistance levels in broilers
##for each year they are present

```{r}
# Creates a table with information on broilers for each year
broiler_information <- Dataset1 %>%
  filter(NORMart == "chi") %>%
  count(quin_res, year) %>%
  mutate(quin_res = ifelse(quin_res == 0,
                           "NonResistant",
                           "Resistant")) %>%
  spread(quin_res, n) %>%
  mutate(n = NonResistant + Resistant,
         pRes = Resistant/n*100) %>%
  group_by(Resistant, n) %>%
  mutate(lwr = get_binCI(Resistant, n)[1],
         upr = get_binCI(Resistant, n)[2]) %>%
  select(year, n, Resistant, pRes, lwr, upr)
broiler_information[,4:6] <- round(broiler_information[,4:6],1)
```

#Occurrence of QREC in broilers with confidence intervals

##This plot presents the data in the table above

```{r}
ggplot(broiler_information,aes(factor(year),pRes))+
  geom_col(width = 0.85)+
  geom_errorbar(aes(ymax = upr, ymin = lwr), 
                width = 0.4, 
                alpha = 0.5)+
  labs(x = "Year",
       y = "Occurrence of QREC (%)")+
  theme(panel.background = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        axis.line = element_line(color = "black", size = 0.1),
        axis.title.x = element_text(color = "#333333"),
        axis.title.y = element_text(color = "#333333"),
        text = element_text(size = 15, family = "Times"))+
  scale_x_discrete(labels = c("2006\nn = 190","2009\nn = 162",
                              "2011\nn = 208","2012\nn = 113",
                              "2014\nn = 205","2016\nn = 181"))
```
\newpage

#Percent resistance per species per pop. density category

##This creates a table where percent relative resistance per species
##per population density category is presented.

```{r}
species_and_pop_density<- Dataset2 %>%
  mutate(quin_res = ifelse(quin_res == 0,
                           "NonResistant",
                           "Resistant")) %>%
  count(NORMart.x, kategori, quin_res) %>%
  spread(quin_res, n, fill = 0) %>%
  mutate(PercentResistant = 
         Resistant / (NonResistant + Resistant)*100) %>%
  mutate(test = map2(Resistant, NonResistant, ~ 
                       binom.test(.x, .x + .y) %>% 
                       tidy())) %>%
  unnest() %>%
  transmute(Species = NORMart.x, Pop.density = kategori, 
            PercentResistant, CI_low = conf.low * 100,
            CI_high = conf.high * 100, 
            TotalSamples = Resistant + NonResistant) %>%
  complete(Species, Pop.density) %>%
  fullnames(.,Species) %>%
  rename("Density" = Pop.density,
         "Resistance(%)"=PercentResistant,
         "95 % CI Low" = CI_low, 
         "95 % CI High" = CI_high,
         "Total Samples" = TotalSamples) 

species_and_pop_density[is.na(species_and_pop_density)] <- 0
species_and_pop_density[,3:5] <- round(species_and_pop_density[,3:5],1)

pander(species_and_pop_density)

```
\newpage

# Percent resistance per year per animal species
```{r}
pRes_year_species <- Dataset1 %>%
  count(NORMart, quin_res, year) %>%
  mutate(quin_res = ifelse(quin_res == 0, "Nonresistant", "Resistant")) %>%
  spread(quin_res, n, fill = 0) %>%
  mutate(n = Resistant + Nonresistant,
         pRes = Resistant/n*100) %>%
  group_by(year) %>%
  mutate(mean_pRes = mean(pRes)) %>%
  ungroup() %>%
  fullnames(., NORMart)


ggplot(pRes_year_species, aes(factor(year), pRes))+
  geom_point(shape = 21, aes(fill = NORMart, size = n))+
  geom_crossbar(aes(factor(year),
                    mean_pRes,
                    ymin = mean_pRes,
                    ymax = mean_pRes),
                fatten = 0.5,
                color = "black")+
  scale_fill_manual(values = c('#a6cee3','#1f78b4','#b2df8a',
                                  '#33a02c','#fb9a99','#e31a1c',
                                  '#fdbf6f','#ff7f00','#cab2d6',
                                  '#6a3d9a','#ffff99'))+
  theme_classic()+
  theme(legend.key.size = unit(5, "point"),
        legend.title = element_blank())+
  labs(x = "Years",
       y = "Percent (%) occurrence of QREC")+
  guides(fill = guide_legend(title = "Animal\nSpecies", override.aes = list(size = 4)),
         size = guide_legend(title = "n"))

```

\newpage

#Statistics

##Chi squared tests for total QREC occurrence each year for all species

```{r}
TotResYearChisq <- Dataset1 %>%
  mutate(quin_res = ifelse(quin_res == 1,
                           "Resistant",
                           "NonResistant")) %>%
  count(year, quin_res) %>%
  spread(quin_res, n, fill = 0)

TotResYearChisq_values <- TotResYearChisq %>%
  slice(expand.grid(1:11, 1:11) %>% rev %>% 
          filter(Var2 < Var1) %>% t) %>%
  mutate(test = rep(1:(n()/2), each = 2)) %>%
  group_by(test) %>%
  do(data_frame(test = first(.$test),
                year1 = first(.$year),
                year2 = last(.$year),
                data = list(matrix(c(.$NonResistant,
                                     .$Resistant),
                                   ncol = 2)))) %>%
  mutate(chi_test = map(data, chisq.test, correct = FALSE)) %>%
  mutate(p.value = map_dbl(chi_test, "p.value")) %>%
  ungroup() %>%
  select(year1, year2, p.value) %>%
  mutate(Significance = ifelse(p.value < 0.05,
                               "Significant",
                               "Not significant")) %>%
  rename("Year One"="year1","Year Two"="year2","p-Value"="p.value")

TotResYearChisq_values[,3] <- round(TotResYearChisq_values[,3],3)

pander(TotResYearChisq_values)

```
\newpage

## Chi squared tests for each species compared to one another

```{r}
#Prepares a table for the chi squared tests
species_table <- Dataset1 %>%
  count(NORMart, quin_res) %>%
  mutate(quin_res = ifelse(quin_res == 0,
                           "NonResistant",
                           "Resistant")) %>%
  spread(quin_res, n, fill = 0)

species_chisq <- species_table %>%
  slice(expand.grid(1:11, 1:11) %>%
          rev %>%
          filter(Var2 < Var1) %>% 
          t) %>%
  mutate(test = rep(1:(n()/2), each = 2)) %>%
  group_by(test) %>%
  do(data_frame(test = first(.$test),
                species1 = first(.$NORMart),
                species2 = last(.$NORMart),
                data = list(matrix(c(.$NonResistant,
                                     .$Resistant),
                                   ncol = 2)))) %>%
  mutate(chi_test = map(data, chisq.test, correct = FALSE)) %>%
  mutate(p.value = map_dbl(chi_test, "p.value")) %>%
  ungroup() %>%
  select(species1, species2, p.value) %>%
  fullnames(., species1, species2) %>%
  mutate(Significance = ifelse(p.value < 0.05,
                               "Significant",
                               "Not significant")) %>%
  rename("Species One"=species1,
         "Species Two"=species2,
         "p-Value"="p.value")

pander(species_chisq)
```
\newpage

## Chi squared test for broilers each year

```{r}
broilers <- Dataset1 %>%
  filter(NORMart == "chi") %>%
  count(NORMart, quin_res, year) %>%
  mutate(quin_res = ifelse(quin_res == 0,
                           "NonResistant",
                           "Resistant")) %>%
  spread(quin_res, n, fill = 0)
  
broiler_chisq <- broilers %>%
  slice(expand.grid(1:6, 1:6) %>% rev %>% 
          filter(Var2 < Var1) %>% t) %>%
  mutate(test = rep(1:(n()/2), each = 2)) %>%
  group_by(test) %>%
  do(data_frame(test = first(.$test),
                year1 = first(.$year),
                year2 = last(.$year),
                data = list(matrix(c(.$NonResistant,
                                     .$Resistant),
                                   ncol = 2)))) %>%
  mutate(chi_test = map(data, chisq.test, correct = FALSE)) %>%
  mutate(p.value = map_dbl(chi_test, "p.value")) %>%
  ungroup() %>%
  select(year1, year2, p.value) %>%
  mutate(Significance = ifelse(p.value < 0.05,
                               "Significant",
                               "Not significant")) %>%
  rename("Year One"="year1",
         "Year Two"="year2",
         "p-Value"="p.value")

pander(broiler_chisq)
```
\newpage

## Chi squared test for broilers and wild birds pooled,
## against the rest of the species

```{r}
chisq_avichi_rest <- matrix(c(1317,3186,45,20), nrow = 2)
chisq_avichi <- chisq.test(chisq_avichi_rest, correct = F)
pander(chisq_avichi)
```

## Chi squared tests on same species for each population density category

```{r}
#Prepares a table for the chi squared tests
category_table <- Dataset2 %>%
  count(NORMart.x, quin_res, kategori) %>%
  mutate(quin_res = ifelse(quin_res == 0,
                           "NonResistant",
                           "Resistant")) %>%
  spread(quin_res, n, fill = 0)

pop_density_species_chisq <- category_table %>%
  group_by(NORMart.x) %>%
  slice(c(1, 2, 1, 3, 2, 3)) %>%
  mutate(test = rep(1:(n()/2), each = 2)) %>%
  group_by(NORMart.x, test) %>%
  do(data_frame(NORMart.x = first(.$NORMart.x),
                test = first(.$test[1]),
                category1 = first(.$kategori),
                category2 = last(.$kategori),
                data = list(matrix(c(.$NonResistant,
                                     .$Resistant),
                                   ncol = 2)))) %>%
  mutate(chi_test = map(data, chisq.test, correct = FALSE)) %>%
  mutate(p.value = map_dbl(chi_test, "p.value")) %>%
  ungroup() %>%
  select(NORMart.x, category1, category2, p.value) %>%
  fullnames(.,NORMart.x) %>%
  mutate(significance = ifelse(p.value < 0.05,
                               "Significant",
                               "Not significant")) %>%
  rename(Species = "NORMart.x",
         "Density One"="category1",
         "Density Two"="category2")

pander(pop_density_species_chisq)
```

## Correlation test for Ciprofloxacin vs. Nalidixic acid

```{r}
cor_test <- cor.test(Dataset1$CIP, Dataset1$NAL)
pander(cor_test)

```

